\thesistitle

### 2025-07-20 manuscript/main.tex
titlepage as per example

\documentclass[11pt]{report}
\usepackage[paper=a4paper,margin=2.5cm]{geometry}
\usepackage[outputdir=out, newfloat]{minted}
\setminted{
  fontsize=\small
}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{
  arrows.meta,      % nicer arrow tips
  shapes.geometric, % rectangle, diamond, etc.
  positioning,      % easy relative placement
  chains,           % flow‐style “chain” syntax
  calc             % coordinate arithmetic
}
\tikzset{
  process/.style = {
    rectangle,
    draw,
    minimum width=3cm,
    minimum height=1cm,
    text centered,
    on chain
  },
  arrow/.style   = {
    -Stealth,
    thick
  },
  start chain    = going below,  % default chain direction
  node distance  = 8mm           % space between chain nodes
}
\usepackage{amsfonts}
\usepackage{fontspec}

% Point fontspec at your local fonts folder:
\setmainfont{EBGaramon}[
  Path      = font/,       % relative to main.tex
  Extension = .ttf,
  UprightFont    = *d12-Regular,   % MyFont-Regular.ttf
  BoldFont       = *d08-Regular,       % MyFont-Bold.ttf
  ItalicFont     = *d12-Italic,
  BoldItalicFont = *d08-Italic
]

\usepackage{forest}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[colorlinks,linkcolor=blue,urlcolor=cyan]{hyperref}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{titlesec}
\titleformat{\chapter}[hang]{\normalfont\huge\bfseries}{}{0pt}{}
\titlespacing*{\chapter}{0pt}{1.5em}{1em}
%=== Metadata ===
\newcommand{\thesistitle}{Your Thesis Title Here}
\newcommand{\authorname}{Rafael Tanzer}
\newcommand{\supervisor}{Michele Chiaria, Simone , Francesco Pontigia and Ezzio Bartocci}
\newcommand{\institution}{TU Vienna}
\newcommand{\department}{Cyber-Physical-Systems, ..., ...}
\newcommand{\logoimage}{graphics/logo.png} % Path to your logo image
\newcommand{\submissionmonth}{May}
\newcommand{\submissionyear}{2025}
%================

\usepackage[backend=biber, style=numeric]{biblatex}
\addbibresource{bibliography/theory.bib}
\addbibresource{bibliography/implementation.bib}
\usepackage{./simplebnf}
\def\code#1{\texttt{\detokenize{#1}}}

\definecolor{lightgray}{gray}{0.97}

@@

\begin{titlepage}
  \centering
  \vspace*{1cm}

  % University logo
  \includegraphics[width=0.25\textwidth]{graphics/tuWienLogo.png}
  \\[1cm]

  % Thesis title
  {\Huge \thesistitle\\[1.5cm]}

  % Project logo under title, smaller size
  \includegraphics[width=0.15\textwidth]{\logoimage}\\[1cm]

  % Author\maketitle
  {\Large \authorname\\[0.5cm]}

  % Department and institution
  {\large \department\\
  \institution\\[1.5cm]}

  % Supervisor
  {\large Supervisor: \supervisor\\[2cm]}

  % Submission date
  {\large \submissionmonth~\submissionyear\\}

  \vfill

  % Bottom of the page
  \vspace*{0.5cm}
  {\small This thesis is submitted in partial fulfillment of the requirements for the degree of \textit{Bachelor of Science}.}

\end{titlepage}

>>>>

% Document class and page setup
\documentclass[11pt]{report}
\usepackage[paper=a4paper,margin=2.5cm]{geometry}

%=== Syntax highlighting ===
\usepackage[outputdir=out, newfloat]{minted}
\setminted{
  fontsize=\small
}

%=== Fonts ===
\usepackage{fontspec}
% Point fontspec at your local fonts folder:
\setmainfont{EBGaramon}[
  Path      = font/,       % relative to main.tex
  Extension = .ttf,
  UprightFont    = *d12-Regular,
  BoldFont       = *d08-Regular,
  ItalicFont     = *d12-Italic,
  BoldItalicFont = *d08-Italic
]

%=== Mathematics ===
\usepackage{amsfonts}

%=== Layout and typography ===
\usepackage{caption}
\usepackage{multicol}
\usepackage{microtype}
\usepackage{titlesec}
\titleformat{\chapter}[hang]{\normalfont\huge\bfseries}{}{0pt}{}
\titlespacing*{\chapter}{0pt}{1.5em}{1em}
% Prevent chapters from starting on a new page
\usepackage{etoolbox}
\makeatletter
  % In the standard report class \chapter does:
  %    \if@openright \cleardoublepage \else \clearpage \fi
  % so we remove that entire conditional.
  \patchcmd{\chapter}
    {\if@openright\cleardoublepage\else\clearpage\fi}
    {}
    {}{}

  % Starred chapters (\chapter*) invoke \@schapter and do a \clearpage
  \patchcmd{\@schapter}
    {\clearpage}
    {}
    {}{}
\makeatother
%=== Graphics and diagrams ===
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{xcolor}
\definecolor{lightgray}{gray}{0.97}

\usepackage{tikz}
\usetikzlibrary{
  arrows.meta,      % nicer arrow tips
  shapes.geometric, % rectangle, diamond, etc.
  positioning,      % easy relative placement
  chains,           % flow‐style “chain” syntax
  calc              % coordinate arithmetic
}
\tikzset{
  process/.style = {
    rectangle,
    draw,
    minimum width=3cm,
    minimum height=1cm,
    text centered,
    on chain
  },
  arrow/.style   = {
    -Stealth,
    thick
  },
  start chain    = going below,  % default chain direction
  node distance  = 8mm            % space between chain nodes
}

\usepackage{forest}

%=== Hyperlinks ===
\usepackage[colorlinks,linkcolor=blue,urlcolor=cyan]{hyperref}

%=== Metadata ===
\newcommand{\thesistitle}{VS Code Language Extension - MiniProb}
\newcommand{\authorname}{Rafael Tanzer}
\newcommand{\supervisor}{Michele Chiaria, Simone , Francesco Pontigia and Ezzio Bartocci}
\newcommand{\institution}{TU Vienna}
\newcommand{\department}{Cyber-Physical-Systems, ..., ...}
\newcommand{\logoimage}{graphics/logo.png} % Path to your logo image
\newcommand{\submissionmonth}{May}
\newcommand{\submissionyear}{2025}
\newcommand{\submissionday}{24}

%=== File Metadata ======
\begin{filecontents*}[overwrite]{\jobname.xmpdata}
\Author{\authorname}
\Title{\thesistitle}
\Language{de-AT}
\Keywords{a\sep list\sep of\sep keywords}
\Publisher{TU Wien}
\Subject{Thesis}
\end{filecontents*}

%=== Bibliography ===
\usepackage[backend=biber, style=numeric]{biblatex}
\addbibresource{bibliography/theory.bib}
\addbibresource{bibliography/implementation.bib}

%=== Custom BNF and code styling ===
\usepackage{./simplebnf}
\def\code#1{\texttt{\detokenize{#1}}}

@@

%—— German Title Page ——
\begin{titlepage}
  % Top‑left TU Wien logo
  \noindent
  \includegraphics[width=0.25\textwidth]{graphics/tuWienLogo.png}

  % push down to center the main block
  \vspace*{\fill}

  % Centered metadata block
  \begin{center}
    {\Huge \thesistitle}\\[1cm]
    \includegraphics[width=0.15\textwidth]{\logoimage}\\[1cm]
    {\LARGE BACHELORARBEIT}\\[0.5cm]
    zur Erlangung des akademischen Grades\\[0.5cm]
    {\Large Bachelor of Science}\\[0.5cm]
    im Rahmen des Studiums\\[0.5cm]
    \textit{Bachelor programme Software \& Information Engineering}\\[1cm]
    eingereicht von\\[0.5cm]
    {\Large \authorname}\\[0.5cm]
    Matrikelnummer 12224207
  \end{center}

  % Left‑aligned faculty / advisors block
  \vspace*{1cm}
  \noindent
  an der Fakultät für Informatik\\
  der Technischen Universität Wien\\

  \vspace*{0.5cm}
  \noindent
  \begin{tabular}{@{}r @{\quad} l@{}}
    Betreuung: & Ezzio Bartocci\\
    Mitwirkung: & Michele Chiari\\
              & Francesco Pontiggia
  \end{tabular}

  % push rest down to bottom
  \vspace*{\fill}

  \noindent
  \makebox[\textwidth]{%
    {\small Vienna, \submissionday~\submissionmonth~\submissionyear}%
    \hfill
    \begin{tabular}[t]{@{\hspace{2cm}}c@{\hspace{0.2cm}}c@{}}
      \rule{5cm}{0.4pt} & \rule{5cm}{0.4pt} \\[-0.3ex]
      {\small \authorname} & {\small Ezzio Bartocci}
    \end{tabular}%
  }
\end{titlepage}

%—— English Title Page ——
\begin{titlepage}
  % Top‑left TU Wien logo
  \noindent
  \includegraphics[width=0.25\textwidth]{graphics/tuWienLogo.png}

  % push down to center the main block
  \vspace*{\fill}

  % Centered metadata block
  \begin{center}
    {\Huge \thesistitle}\\[1cm]
    \includegraphics[width=0.15\textwidth]{\logoimage}\\[1cm]
    {\LARGE BACHELOR'S THESIS}\\[0.5cm]
    for the degree of\\[0.5cm]
    {\Large Bachelor of Science}\\[0.5cm]
    within the study programme of\\[0.5cm]
    \textit{Bachelor programme Software \& Information Engineering}\\[1cm]
    submitted by\\[0.5cm]
    {\Large \authorname}\\[0.5cm]
    Student ID 12224207
  \end{center}

  % Left‑aligned faculty / advisors block
  \vspace*{1cm}
  \noindent
  to the Faculty of Informatics\\
  at the TU Wien\\

  \vspace*{0.5cm}
  \noindent
  \begin{tabular}{@{}r @{\quad} l@{}}
    Advisor:    & Ezio Bartocci\\
    Assistance: & Michele Chiari\\
              & Francesco Pontiggia
  \end{tabular}

  % push rest down to bottom
  \vspace*{\fill}

  \noindent
  \makebox[\textwidth]{%
    {\small Vienna, \submissionday~\submissionmonth~\submissionyear}%
    \hfill
    \begin{tabular}[t]{@{\hspace{2cm}}c@{\hspace{0.2cm}}c@{}}
      \rule{5cm}{0.4pt} & \rule{5cm}{0.4pt} \\[-0.3ex]
      {\small \authorname} & {\small Ezzio Bartocci}
    \end{tabular}%
  }
\end{titlepage}

### 2025-07-20 manuscript/part1.tex
titlepage as per example

While BNF provides a clear, formal way to describe language syntax, it suffers from verbosity and redundancy - pure BNF grammars often become bloated when encoding
common patterns like optionals or repetitions requiring auxiliary nonterminals for a sufficient description and, over time, has been extended and modified for various
use-cases spawning a new family of grammar notations. The \textbf{Extended Backus-Naur Form} extends BNF by adding more expressive meta-syntax, introducing operators
similar to \textit{Regular Expressions} allowing the use of optional or repeatable expressions, commonly indicated by brackets but not limited to '[\dots]' and '\{\dots\}',
and the use of comments. Such extensions make grammars more compact and easier to read without changing the fundamental class of languages they describe.
Multiple distinct instances of EBNF's have been development, all with minor syntactic differences, yet there really is no on-for-all EBNF used in practice, although
a standardized ISO/IEC 14977 version exists. \cite{jinks2004bnf,jinks2004ebnfvariants}\\
Within the scope of my thesis, \textit{?NAME?}, Extended Backus-Naur Form (EBNF) is particularly significant, since Langium employs its own custom EBNF dialect ?makeRefTo?
to define the grammars underpinning its language-assistance features. \ref{sec:langium-grammar}

@@

\subsubsection*{Lexers}
todo

@@

Lastly, MiniProb also includes a few more constructs which are omitted here as, they are currently not covered by the language extension.
These constructs include \code{probabilistic query: ...} defining the type of query, \code{formulas = ...}: specified by POTL formalisms and the definition of \emph{modules} \cite{guideMiniProb}.

>>>>

While BNF provides a clear, formal way to describe language syntax, it suffers from verbosity and redundancy - pure BNF grammars often become bloated when encoding
common patterns like optionals or repetitions requiring auxiliary nonterminals for a sufficient description and, over time, has been extended and modified for various
use-cases spawning a new family of grammar notations. The \textbf{Extended Backus-Naur Form} extends BNF by adding more expressive meta-syntax, introducing operators
similar to \textit{Regular Expressions} allowing the use of optional or repeatable expressions, commonly indicated by brackets but not limited to '[\dots]' and '\{\dots\}',
and the use of comments. Such extensions make grammars more compact and easier to read without changing the fundamental class of languages they describe.
Multiple distinct instances of EBNF's have been development, all with minor syntactic differences, yet there really is no on-for-all EBNF used in practice, although
a standardized ISO/IEC 14977 version exists. \cite{jinks2004bnf,jinks2004ebnfvariants}\\
Within the scope of this thesis, the Extended Backus-Naur Form (EBNF) is particularly significant, since Langium employs its own custom EBNF dialect \ref{sec:langium-grammar}
to define the grammars underpinning its language-assistance features. \ref{sec:langium-grammar}

@@

@@

Lastly, MiniProb also includes a few more constructs which are omitted here as, they are currently not covered by the language extension.
These constructs include \code{probabilistic query: ...} defining the type of query, \code{formulas = ...} - specified by POTL formalisms - and the definition of \emph{modules} \cite{guideMiniProb}.
